\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{fullpage}
\newcommand{\dsum}{\displaystyle\sum}
\newcommand{\dbcup}{\displaystyle\bigcup}
\newcommand{\dbcap}{\displaystyle\bigcap}
\newcommand{\dcup}{\displaystyle\cup}
\newcommand{\dcap}{\displaystyle\cap}
\newcommand{\Pb}{\mathbb{P}}
\newcommand{\bkt}[1]{\left(#1\right)}
\title{MA2040: Probability, Statistics and Stochastic Processes\\
Problem Set-I}
\author{Sivaram Ambikasaran}
\begin{document}
	\maketitle
	\begin{enumerate}
		\item
		A six-sided die is made in a way that each even face is twice as likely as each odd face. All even faces are equally likely, as are all odd faces. Construct a probabilistic model for a single roll of this die and find the probability that the outcome is less than 4.
		\item
		Let $S_1,S_2,\ldots,S_n$ be a partition of the sample space $\Omega$.
		\begin{enumerate}
			\item
			Show that for any event $A$,
			$$\Pb(A) = \dsum_{i=1}^n \Pb\bkt{A \dcap S_i}$$
			\item
			Use the previous part to show that, for events $A$, $B$ and $C$,
			$$\Pb\bkt{A} = \Pb\bkt{A \dcap B} + \Pb\bkt{A \dcap C} + \Pb\bkt{A \dcap B^c \dcap C^c} - \Pb\bkt{A \dcap B \dcap C}$$
		\end{enumerate}
		\item
		\begin{enumerate}
			\item
			Prove that for any two events $A$ and $B$, we have
			$$\Pb\bkt{A \dcap B} \geq \Pb(A) + \Pb(B) - 1$$
			\item
			Using the above, establish the following generalization:
			$$\Pb\bkt{A_1 \dcap A_2 \dcap \cdots \dcap A_n} \geq \Pb\bkt{A_1} + \Pb\bkt{A_2} + \cdots + \Pb\bkt{A_n} - \bkt{n-1}$$
		\end{enumerate}
		\item
		Let $\Omega = \{\bkt{x_1,x_2,x_3}: x_1+x_2+x_3 = 2019\}$ and $x_1,x_2,x_3$ are positive integers. Assuming all the elements in $\Omega$ are equally likely to be drawn, what is the probability of having $x_1,x_2$ and $x_3$ (all three) to be odd?
		\item
		Two fair $6$-sided dice are rolled.
		\begin{enumerate}
			\item
			Given that the roll results in a sum of $4$ or less, find the conditional probability that both dice show the same number.
			\item
			Given that the two dice land on different numbers, find the conditional probability that at least one die roll is a $6$.
		\end{enumerate}
		\item
		A batch of $100$ items is inspected by testing $4$ randomly selected items. If one of the four is defective, the batch is rejected. What is the probability that the batch is accepted, if it contains exactly five defectives?
		\item
		Consider a coin that comes up heads with probability $p$ and tails with probability $1-p$. Let $q_n$ be the probability of obtaining even number of heads in $n$ independent tosses. Derive a recursion that relates $q_n$ to $q_{n-1}$ and establish the formula
		$$q_n = \dfrac{1+\bkt{1-2p}^n}2$$
		\item
		Two playes $X$ and $Y$ alternately roll a pair of unbiased dice. $X$ wins if on a throw he gets a sum of $6$ before $Y$ gets a sum of $7$; $Y$ wins if he obtains a sum of $7$ before $X$ obtains a sum of $6$; If $X$ begins the game, prove that his probability of winning is $30/61$.
		\item
		In a deck of cards, let $A$ be the event of drawing a spade and $B$ be the event of drawing a king. Assuming that all cards are equally likely to be drawn, obtain $\Pb(A \vert B)$, $\Pb (B \vert A)$. Are $A$ and $B$ independent?
		\item
		Let $p_X$ denote the probability that India will play a team $X$ in the World Cup Final $2019$, given that India will qualify for the World Cup Final $2019$. Let $q_X$ be the probability of Kohli scoring a century against team $X$.

		\begin{table}[!htbp]
			\begin{center}
			\begin{tabular}{|c|c|c|}
				\hline
				Teams & $p_X$ & $q_X$\\
				\hline
				\hline
				Afghanistan & $0.02$ & $0.50$\\
				\hline
				Australia & $0.10$ & $0.30$ \\
				\hline
				Bangladesh & $0.03$ & $0.40$\\
				\hline
				England & $0.30$ & $0.10$\\
				\hline
				New Zealand & $0.15$ & $0.25$\\
				\hline
				Pakistan & $0.10$ & $0.15$\\
				\hline
				South Africa & $0.20$ & $0.20$\\
				\hline
				SriLanka & $0.05$ & $0.25$\\
				\hline
				West Indies & $0.05$ & $0.20$\\
				\hline
			\end{tabular}
			\end{center}
		\end{table}
		\begin{enumerate}
			\item
			From the information given above, find the probability that Kohli will score a century in the World Cup Final $2019$. 
			\item
			Given that Kohli scores a century in the World Cup Final $2019$, which team is
			\begin{enumerate} 
				\item Most likely to have played the Final along with India
				\item Least likely to have played the Final along with India
			\end{enumerate}
		\end{enumerate}
		\item
		A test for certain rare virus correctly predicts that the person has a virus $99\%$ of the time and correctly identifies that the person doesn't carry a virus $98\%$ of the time. It is known $1\%$ of the population carries the virus. What is the probability of a person actually having the disease, if he has tested positive to the test?
	\end{enumerate}
\end{document}