\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{tikz}

\newcommand{\dsum}{\displaystyle\sum}
\newcommand{\dint}{\displaystyle\int}
\newcommand{\abs}[1]{\displaystyle\left\lvert#1\right\rvert}
\newcommand{\dbcup}{\displaystyle\bigcup}
\newcommand{\dbcap}{\displaystyle\bigcap}
\newcommand{\dcup}{\displaystyle\cup}
\newcommand{\dcap}{\displaystyle\cap}
\newcommand{\Pb}{\mathbb{P}}
\newcommand{\Eb}{\mathbb{E}}
\newcommand{\bkt}[1]{\left(#1\right)}
\title{MA2040: Probability, Statistics and Stochastic Processes\\
Problem Set-IV}
\author{Sivaram Ambikasaran}
\begin{document}
	\maketitle
	\begin{enumerate}
		\item
		Let $X$ and $Y$ have joint pdf $f_{X,Y}(x,y) = \begin{cases}
		6 e^{-\bkt{2x+3y}} & x,y \geq 0\\
		0 & \text{otherwise}
		\end{cases}$
	\begin{itemize}
		\item
		Are $X$ and $Y$ independent?
		\item
		Find $\mathbb{E}\bkt{Y \mid X>2}$
		\item
		Find $\Pb\bkt{X > Y}$
	\end{itemize}
	\item
	Let $X \sim \text{Uniform}(1,2)$ and given $X=x$, $Y$ follows an exponential distribution with $Y \sim \text{EXP}\bkt{\lambda=x}$. Find the covariance of $X$ and $Y$.
	\item
	Let $X$ and $Y$ be independent standard normal random variables. Find covariance of $Z$ and $W$, where $Z = 1+X+XY^2$ and $W = 1+X$.
	\item
	A fair die is rolled $n$ times. Let $X$ be the number of $1$'s and $Y$ be the number of $2$'s. Find the correlation coefficient $\rho(X,Y)$.
	\item
	Let $32$ people sit around a round table. Each person tosses a fair coin. Anyone whose outcome is different from both his neighbors is taken out from the group. If $X$ is the total number of such persons, find variance of $X$.
	\item
	The moment generating function of a random variable $X$ is given by
	$$M_X(s) = \dfrac2{2-s}, \,\,\, \forall s \in (-\infty,2)$$
	Find the distribution of $X$.
	\item
	Let $X \sim \text{Binomial}(n,p)$ and $Y \sim \text{Binomial}(m,p)$ be independent random variables. Show that $X+Y \sim \text{Binomial}(m+n,p)$.
	\item
	Show that the function $e^{-s}$ is a moment generating function.
	\item
	Show that if $M(s)$ is a moment generating function, so is $M(cs)$.
	\item
	Show that if $M(s)$ is a moment generating function, $cM(s)$ cannot be a moment generating function for $c \neq 1$.
	\item
	Show that if $M(s)$ is a moment generating function, then $e^{-s}M(s)$ is also a moment generating function.
	\item
	The moment generating function of a random variable $X$ is
	$$M_X(s) = \dfrac{e^{-2s}}6 + \dfrac{e^{-s}}3 + \dfrac{e^{s}}4 + \dfrac{e^{2s}}4$$
	Show that $P\bkt{\abs{X} \leq 1} = \dfrac7{12}$.
	\end{enumerate}
\end{document}